---
title: "moco-3-pattern-adult-1F1.Rmd"
author: "rogilmore"
date: "`r Sys.Date()`"
output: html_document
---

This document demonstrates the analyses reported in Gilmore, Thomas, and Fesi (2015).

## Set-up

### Harmonic of interest and participant group, alpha level

```{r settings}
this.harm <- "1F1"
p.thresh = 0.0005
this.group <- 'adult'

dir.figs <- 'figs'
fn.data <- paste("moco-3-pattern-", this.group, ".csv", sep="")
fn.egi <- "egi.csv"
```

### Source libraries and load data.

Source libraries

```{r libraries, warning=FALSE, message=FALSE}
# Libraries
library(ggplot2)
library(dplyr)
library(png)
library(gridExtra)
library(tidyr)
library(knitr)
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path=paste(dir.figs, "/", sep=""), dpi=300)
```

Load MOCO dataframe.

```{r load-data}
df.moco <- read.csv(fn.data)
```

### EGI channels

Load EGI channel positions.

```{r load-egi-channel-location-data}
df.egi <- read.csv(fn.egi)

# Load topo ears & nose from file
img <- readPNG(paste(dir.figs, "topoplot.png", sep="/"))

# Add a raster
topo_ears_nose <- rasterGrob(img, interpolate=TRUE)
```

## `r this.harm` mass univariate analysis

Create mass univariate function.

```{r mass-univariate-function}
mass.univariate <- function( ch, df, harm, method ) { 

  #Select data frame
  local.df = df[ df$Channel == ch & df$Harm == harm,]

  #Select method and conduct analysis
  if (method =="manova.ampl.complex"){
    this.stats <- manova( formula = cbind(Sr, Si) ~ Pattern*Speed + Error(iSess), data=local.df )   } else {
     this.stats = NA
   }
   return( this.stats )
}
```

### Mass univariate MANOVA on `r this.harm`

MANOVA on `r this.harm` with Pattern, Speed, and Pattern x Speed. Analyze and write write output to file.

```{r mass-univariate-manova}
maov.list = lapply( unique( df.moco$Channel ), mass.univariate, df=df.moco, harm=this.harm, method="manova.ampl.complex")
maov.summ = lapply( maov.list, summary )
```

#### Extract channel-level statistics

Extract statistics, $F$ and $p$ values, from MANOVA summary. Create data frame for plotting and tweak factor levels and labels for aesthetic reasons.

```{r extract-statistics}
#'Error: Within.stats17' denominator df for Pattern
#'Error: Within.stats18' denominator df for Speed
#'Error: Within.stats19' denominator df for Patt*Spd
#'Error: Within.stats13' num df for Pattern
#'Error: Within.stats14' num df for Speed
#'Error: Within.stats15' num df for Patt*Spd
#'Error: Within.stats9'  F for Pattern
#'Error: Within.stats10' F for Speed
#'Error: Within.stats11' F for Patt*Spd
#'Error: Within.stats21' p for Pattern
#'Error: Within.stats22' p for Speed
#'Error: Within.stats23' p for Patt*Spd

# Function to extract stats
moco.stats <- function(model.list) {
  ml = unlist(model.list)
  new.vals <- as.numeric( c( ml['Error: Within.stats9'], 
                             ml['Error: Within.stats10'], 
                             ml['Error: Within.stats11'], 
                             ml['Error: Within.stats13'], 
                             ml['Error: Within.stats14'], 
                             ml['Error: Within.stats15'],
                             ml['Error: Within.stats17'], 
                             ml['Error: Within.stats18'], 
                             ml['Error: Within.stats19'], 
                             ml['Error: Within.stats21'], 
                             ml['Error: Within.stats22'], 
                             ml['Error: Within.stats23']) )
  names(new.vals) <- c("F-Pattern", "F-Speed", "F-Patt*Spd", "df-Pattern", "df-Speed", "df-Patt*Spd", "df-Pattern-denom", "df-Speed-denom", "df-Patt*Spd-denom", "p-Pattern", "p-Speed", "p-Patt*Spd")
  return( rbind( new.vals ) )
}

# Extract stats and assemble into data frame
list.stats <- data.frame( t( sapply(maov.summ, moco.stats) ) )
names( list.stats ) <- rep( c("Pattern", "Speed", "Patt_Spd"), 4)
list.F <- list.stats[,1:3]
list.dfNum <- list.stats[,4:6]
list.dfDen <- list.stats[,7:9]
list.dfp <- list.stats[,10:12]
F_val <- data.frame(list.F) %>% gather(Effect, F_val, Pattern:Patt_Spd)
dfNum <- data.frame(list.dfNum) %>% gather(Effect, dfNum, Pattern:Patt_Spd)
dfDen <- data.frame(list.dfDen) %>% gather(Effect, dfDen, Pattern:Patt_Spd)
pvals <- data.frame(list.dfp) %>% gather(Pvals, pvals, Pattern:Patt_Spd)
df.stats <- data.frame(Chan = rep( 1:128, 3), 
                       Cond = rep(c("Pattern", "Speed", "Patt*Spd"), 
                                  c(128, 128, 128)),
                       Fvals = F_val[,2],
                       FdfNum = dfNum[,2],
                       FdfDen = dfDen[,2],
                       Pvals = pvals[,2],
                       xpos=rep(df.egi$xpos,3),
                       ypos=rep(df.egi$ypos,3))

# Cut pvals
pvals.cuts = c(-.01,.0001, .0005, .001, .005, .01, .05, 1)
pvals.lbls = c("<.0001","<.0005", "<.001", "<.005", "<.01", "<.05", "ns")

# Create cuts based on p-value levels
Pvals.cuts = cut( df.stats$Pvals, breaks=pvals.cuts, labels=pvals.lbls)
Pvals.cuts = ordered( Pvals.cuts, levels = rev( pvals.lbls ) )
df.stats$Pvals.cuts = Pvals.cuts

df.stats$Cond = ordered( df.stats$Cond, levels=c("Pattern", "Speed", "Patt*Spd"))
```

#### Figure -- `r paste(this.harm, " Effects by Channel", sep="")`

```{r figure-1-1F1}
# Delete scales from plots
yquiet = scale_y_continuous("", breaks=NULL)
xquiet = scale_x_continuous("", breaks=NULL)

# Plot theme for channel topo
pl.theme.topo <- theme( plot.title = element_text(lineheight=.8, face ="bold", vjust=2, size=16),
         legend.title=element_text(size=0),
         legend.text=element_text(size=12, face="bold"),
         legend.position="bottom"
         )

# Plot
pl <- ggplot(data=df.stats, aes(x=xpos, y=ypos, color=Pvals.cuts, size=Pvals.cuts)) +
  geom_point() +
  facet_grid(facets = . ~ Cond) +
  coord_fixed() + 
  xquiet + 
  yquiet + 
  pl.theme.topo

# Add ears, nose
pl + annotation_custom(topo_ears_nose, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)
```

#### Table of `r this.harm` Channel Statistics for Pattern

```{r table-1}
df.stats %>% 
  filter( Cond == "Pattern", Pvals < p.thresh  ) %>% 
  select( Chan, Fvals, FdfNum, FdfDen, Pvals ) %>% 
  kable()
```

#### Table of `r this.harm` Channel Statistics for Speed

```{r table-2}
df.stats %>% 
  filter( Cond == "Speed", Pvals < p.thresh  ) %>% 
  select( Chan, Fvals, FdfNum, FdfDen, Pvals ) %>% 
  kable()
```

#### Extract data for channels below threshold

```{r}
this.cond <- "Pattern"
chans.below.thresh <- df.stats %>%
  filter( Cond == this.cond, Pvals <= p.thresh ) %>%
  select( Chan )

chans.below.thresh <- chans.below.thresh$Chan

df.below.thresh <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm )

n.chans.below <- dim(df.below.thresh)[1]
```

There were `r n.chans.below` channels that met criterion for `r this.cond`.

#### Figure -- `r paste( this.harm, " Channels Below p<.0005 for ", this.cond, sep="")`

Calculate mean values for each participant, by channel and pattern for the real (Sr) and imaginary (Si) components at the selected harmonic. Then, calculate the group means for Sr and Si and the group standard deviations and root-mean-square (RMS) amplitudes (and standard errors) from the group mean Sr and Si values.

Then plot.

```{r figure-2-1F1}
if (n.chans.below) {
  df.below.thresh.pattern <- df.below.thresh %>%
    group_by( Channel, Pattern, iSess ) %>%
    summarise( sr.sub.mean=mean(Sr), 
               si.sub.mean=mean(Si),
               rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
    group_by( Channel, Pattern ) %>%
    summarise( sr.group.mean=mean( sr.sub.mean ),
               si.group.mean=mean( si.sub.mean ),
               sr.group.sd=sd(sr.sub.mean),
               si.group.sd=sd(si.sub.mean),
               rms.amp=sqrt( sr.group.mean^2 + si.group.mean^2),
               rms.amp.sem = mean( rms.amp.sub )/sqrt(n()),
               nsubs=n(),
               sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
               si.group.sem=mean(si.sub.mean)/sqrt(n())
    )
  
  limits <- aes( ymax = rms.amp + rms.amp.sem, ymin = rms.amp )
  dodge <- position_dodge( width=0.8 )
  
  pl.theme.bar <- theme(plot.title = element_text(lineheight=.8, face ="bold", vjust=2, size = 20),
                        panel.background = element_rect(fill=NA),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        panel.border = element_rect(fill=NA,color="black", size=.8,
                                                    linetype="solid"),
                        axis.title.x=element_text(vjust=-.6, size=18),
                        axis.title.y=element_text(face="bold",vjust=1, size=18),
                        axis.text=element_text(color="black", size=8),
                        legend.title=element_blank(),
                        legend.text=element_text(size=16), 
                        legend.position="bottom",
                        legend.background=element_blank())
  
  pl.pattern <- ggplot( data=df.below.thresh.pattern ) +
    aes( x=as.factor(Channel), y=rms.amp, fill=Pattern) + 
    geom_bar( stat="identity", width=0.8, position=dodge ) +
    geom_errorbar( limits, position=dodge, width=0.15 ) +
    xlab("Channel") +
    ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
    pl.theme.bar
  pl.pattern
}
```

#### Extract data for channels below threshold

```{r}
this.cond <- "Speed"
chans.below.thresh <- df.stats %>%
  filter( Cond == this.cond, Pvals <= p.thresh ) %>%
  select( Chan )

chans.below.thresh <- chans.below.thresh$Chan

df.below.thresh <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm )

n.chans.below <- dim(df.below.thresh)[1]
```

There were `r n.chans.below` channels that met criterion for `r this.cond`.

#### Figure -- `r paste( this.harm, " Channels Below p<.0005 for ", this.cond, sep="")`

Calculate mean values for each participant, by channel and pattern for the real (Sr) and imaginary (Si) components at the selected harmonic. Then, calculate the group means for Sr and Si and the group standard deviations and root-mean-square (RMS) amplitudes (and standard errors) from the group mean Sr and Si values.

Then plot.

```{r figure-3-1F1}
if (n.chans.below) {
  df.below.thresh.speed <- df.below.thresh %>%
    group_by( Channel, Speed, iSess ) %>%
    summarise( sr.sub.mean=mean(Sr), 
               si.sub.mean=mean(Si),
               rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
    group_by( Channel, Speed ) %>%
    summarise( sr.group.mean=mean( sr.sub.mean ),
               si.group.mean=mean( si.sub.mean ),
               sr.group.sd=sd(sr.sub.mean),
               si.group.sd=sd(si.sub.mean),
               rms.amp=sqrt( sr.group.mean^2 + si.group.mean^2),
               rms.amp.sem = mean( rms.amp.sub )/sqrt(n()),
               nsubs=n(),
               sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
               si.group.sem=mean(si.sub.mean)/sqrt(n())
    )
  
  limits <- aes( ymax = rms.amp + rms.amp.sem, ymin = rms.amp )
  dodge <- position_dodge( width=0.8 )
  
  pl.theme.bar <- theme(plot.title = element_text(lineheight=.8, face ="bold", vjust=2, size = 20),
                        panel.background = element_rect(fill=NA),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        panel.border = element_rect(fill=NA,color="black", size=.8,
                                                    linetype="solid"),
                        axis.title.x=element_text(vjust=-.6, size=18),
                        axis.title.y=element_text(face="bold",vjust=1, size=18),
                        axis.text=element_text(color="black", size=8),
                        legend.title=element_blank(),
                        legend.text=element_text(size=16), 
                        legend.position="bottom",
                        legend.background=element_blank())
  
  pl.pattern <- ggplot( data=df.below.thresh.speed ) +
    aes( x=as.factor(Channel), y=rms.amp, fill=Speed) + 
    geom_bar( stat="identity", width=0.8, position=dodge ) +
    geom_errorbar( limits, position=dodge, width=0.15 ) +
    xlab("Channel") +
    ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
    pl.theme.bar
  pl.pattern
}
```
