---
title: "moco-3-pattern-adult.Rmd"
author: "rogilmore"
date: "`r Sys.Date()`"
output: html_document
---

This document demonstrates the analyses reported in Gilmore, Thomas, and Fesi (2015).

## Document-wide chunk settings

## Setup file paths, load libraries and data.
In a typical set-up, we have the following file directory structure:

project-name/R       # home for the *.RProj, *.R, *.Rmd, and *.html files.
project-name/data    # *.csv files
project-name/figs    # *.pdf or *.png files

```{r file-paths}
# File paths
dir.figs <- "../figs"
dir.data <- "../data"
fn.data <- "moco-3-pattern-adult.csv"
fn.egi <- "egi.csv"

# knitr options to save figures to dir.figs and create 12x6 figures at 300 DPI
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path=paste(dir.figs, "/", sep=""))
```

Source libraries

```{r libraries, warning=FALSE, message=FALSE}
# Libraries
library(ggplot2)
library(dplyr)
library(png)
library(gridExtra)

# Significance threshold
p.thresh = 0.0005
```

Load MOCO dataframe.

```{r load-data}
df.moco <- read.csv(paste(dir.data, fn.data, sep="/"))
```

Analysis proceeds by analyzing the results of the low-order integer harmonics of the global motion coherence modulation (on/off at 1.2 Hz) separately then the 1F2 or dot update rate (24 Hz).

## EGI channels

Load EGI channel positions.

```{r load-egi-channel-location-data}
df.egi <- read.csv(paste(dir.data, fn.egi, sep="/"))
```

#### Figure - EGI Channels

```{r figure-egi-channel-locations}
# Load topo ears & nose from file
img <- readPNG(paste(dir.figs, "topoplot.png", sep="/"))

# Add a raster
topo_ears_nose <- rasterGrob(img, interpolate=TRUE)

# separate frame for annotations

featured <- c(62,75,106,107)
df.annotate <- data.frame(chan=featured, xpos=df.egi$xpos[featured], ypos=df.egi$ypos[featured], Harmonic=c("1F2", "1F2", "1F1", "1F1"))

pl.egi <- ggplot( data=df.egi, aes(x=xpos, y=ypos, label=chan)) +
  geom_text(size=4) +
  coord_fixed() + # 1:1 aspect ratio
  scale_y_continuous("", breaks=NULL) + # omit y axis
  scale_x_continuous("", breaks=NULL) + # omit x axis
  geom_point(data=df.annotate, aes(x=xpos, y=ypos, shape=Harmonic), size=12) +
  scale_shape(solid = FALSE) + # open shapes
  annotation_custom(topo_ears_nose, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  theme( panel.background = element_rect(fill=NA),
         legend.position = "none")  # blank background and no legend
pl.egi
```

## Mass univariate

Create mass univariate function.

```{r mass-univariate-function}
mass.univariate <- function( ch, df, harm, method ) { 

  #Select data frame
  local.df = df[ df$Channel == ch & df$Harm == harm,]

  #Select method and conduct analysis
  if (method =="manova.ampl.complex"){
    this.stats <- manova( formula = cbind(Sr, Si) ~ Pattern*Speed + Error(iSess), data=local.df )   } else {
     this.stats = NA
   }
   return( this.stats )
}
```


```{r}
this.harm = "1F1"
```

### Mass univariate MANOVA on `r this.harm`

MANOVA on `r this.harm` with Pattern, Speed, and Pattern x Speed. Analyze and write write output to file.

```{r 1F1-mass-univariate-manova}
maov.list = lapply( unique( df.moco$Channel ), mass.univariate, df=df.moco, harm=this.harm, method="manova.ampl.complex")
maov.summ = lapply( maov.list, summary )
```

#### Extract channel-level statistics

Extract statistics, $F$ and $p$ values, from MANOVA summary. Create data frame for plotting and tweak factor levels and labels for aesthetic reasons.

```{r 1F1-extract-statistics}
#'Error: Within.stats17' denominator df for Pattern
#'Error: Within.stats18' denominator df for Speed
#'Error: Within.stats19' denominator df for Patt*Spd
#'Error: Within.stats13' num df for Pattern
#'Error: Within.stats14' num df for Speed
#'Error: Within.stats15' num df for Patt*Spd
#'Error: Within.stats9'  F for Pattern
#'Error: Within.stats10' F for Speed
#'Error: Within.stats11' F for Patt*Spd
#'Error: Within.stats21' p for Pattern
#'Error: Within.stats22' p for Speed
#'Error: Within.stats23' p for Patt*Spd

# Function to extract stats
moco.stats <- function(model.list) {
  ml = unlist(model.list)
  new.vals <- as.numeric( c( ml['Error: Within.stats9'], 
                             ml['Error: Within.stats10'], 
                             ml['Error: Within.stats11'], 
                             ml['Error: Within.stats13'], 
                             ml['Error: Within.stats14'], 
                             ml['Error: Within.stats15'],
                             ml['Error: Within.stats17'], 
                             ml['Error: Within.stats18'], 
                             ml['Error: Within.stats19'], 
                             ml['Error: Within.stats21'], 
                             ml['Error: Within.stats22'], 
                             ml['Error: Within.stats23']) )
  names(new.vals) <- c("F-Pattern", "F-Speed", "F-Patt*Spd", "df-Pattern", "df-Speed", "df-Patt*Spd", "df-Pattern-denom", "df-Speed-denom", "df-Patt*Spd-denom", "p-Pattern", "p-Speed", "p-Patt*Spd")
  return( rbind( new.vals ) )
}

# Extract stats and assemble into data frame
list.stats <- data.frame( t( sapply(maov.summ, moco.stats) ) )
names( list.stats ) <- rep( c("Pattern", "Speed", "Patt_Spd"), 4)
list.F <- list.stats[,1:3]
list.dfNum <- list.stats[,4:6]
list.dfDen <- list.stats[,7:9]
list.dfp <- list.stats[,10:12]
F_val <- data.frame(list.F) %>% gather(Effect, F_val, Pattern:Patt_Spd)
dfNum <- data.frame(list.dfNum) %>% gather(Effect, dfNum, Pattern:Patt_Spd)
dfDen <- data.frame(list.dfDen) %>% gather(Effect, dfDen, Pattern:Patt_Spd)
pvals <- data.frame(list.dfp) %>% gather(Pvals, pvals, Pattern:Patt_Spd)
df.stats <- data.frame(Chan = rep( 1:128, 3), 
                       Cond = rep(c("Pattern", "Speed", "Patt*Spd"), 
                                  c(128, 128, 128)),
                       Fvals = F_val[,2],
                       FdfNum = dfNum[,2],
                       FdfDen = dfDen[,2],
                       Pvals = pvals[,2],
                       xpos=rep(df.egi$xpos,3),
                       ypos=rep(df.egi$ypos,3))

# Cut pvals
pvals.cuts = c(-.01,.0001, .0005, .001, .005, .01, .05, 1)
pvals.lbls = c("<.0001","<.0005", "<.001", "<.005", "<.01", "<.05", "ns")

# Create cuts based on p-value levels
Pvals.cuts = cut( df.stats$Pvals, breaks=pvals.cuts, labels=pvals.lbls)
Pvals.cuts = ordered( Pvals.cuts, levels = rev( pvals.lbls ) )
df.stats$Pvals.cuts = Pvals.cuts

df.stats$Cond = ordered( df.stats$Cond, levels=c("Pattern", "Speed", "Patt*Spd"))
```

#### Figure -- `r paste(this.harm, " Effects by Channel", sep="")`

```{r figure-1F1-effects-by-channel}
# Delete scales from plots
yquiet = scale_y_continuous("", breaks=NULL)
xquiet = scale_x_continuous("", breaks=NULL)

# Plot theme for channel topo
pl.theme.topo <- theme( plot.title = element_text(lineheight=.8, face ="bold", vjust=2, size=16),
         legend.title=element_text(size=0),
         legend.text=element_text(size=12, face="bold"),
         legend.position="bottom"
         )

# Plot
pl <- ggplot(data=df.stats, aes(x=xpos, y=ypos, color=Pvals.cuts, size=Pvals.cuts)) +
  geom_point() +
  facet_grid(facets = . ~ Cond) +
  coord_fixed() + 
  xquiet + 
  yquiet + 
  pl.theme.topo

# Add ears, nose
pl + annotation_custom(topo_ears_nose, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)
```

Notice widespread cluster of channels for Pattern, midline cluster for speed..

#### Table of `r this.harm` Channel Statistics for Pattern

```{r table-1F1-channel-statistics-pattern}
df.stats %>% 
  filter( Cond == "Pattern", Pvals < p.thresh  ) %>% 
  select( Chan, Fvals, FdfNum, FdfDen, Pvals ) %>% 
  kable()
```

#### Table of `r this.harm` Channel Satistics for Speed

```{r table-1F1-channel-statistics-speed}
df.stats %>% 
  filter( Cond == "Speed", Pvals < p.thresh  ) %>% 
  select( Chan, Fvals, FdfNum, FdfDen, Pvals ) %>% 
  kable()
```

#### Extract data for channels below threshold

```{r}
chans.below.thresh = df.stats$Chan[df.stats$Pvals <= p.thresh ]

df.below.thresh <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm )
```

#### Figure -- `r paste( this.harm, " Channels Below p<.0005 for Pattern", sep="")`

Calculate mean values for each participant, by channel and pattern for the real (Sr) and imaginary (Si) components at the selected harmonic. Then, calculate the group means for Sr and Si and the group standard deviations and root-mean-square (RMS) amplitudes (and standard errors) from the group mean Sr and Si values.

```{r}
chans.below <- unique( select(filter(df.stats, Cond == "Pattern", Pvals < p.thresh), Chan))

df.below.thresh.pattern <- df.moco %>%
  filter( Channel %in% chans.below, Harm == this.harm  ) %>%
  group_by( Channel, Pattern, iSess ) %>%
  summarise( sr.sub.mean=mean(Sr), 
             si.sub.mean=mean(Si),
             rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
  group_by( Channel, Pattern ) %>%
  summarise( sr.group.mean=mean( sr.sub.mean ),
             si.group.mean=mean( si.sub.mean ),
             sr.group.sd=sd(sr.sub.mean),
             si.group.sd=sd(si.sub.mean),
             rms.amp=sqrt( sr.group.mean^2 + si.group.mean^2),
             rms.amp.sem = mean( rms.amp.sub )/sqrt(n()),
             nsubs=n(),
             sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
             si.group.sem=mean(si.sub.mean)/sqrt(n())
             )
```

Then plot.

```{r figure-1F1-pattern-ampl-adult}
limits <- aes( ymax = rms.amp + rms.amp.sem, ymin = rms.amp )
dodge <- position_dodge( width=0.8 )

pl.theme.bar <- theme(plot.title = element_text(lineheight=.8, face ="bold", vjust=2, size = 20),
                    panel.background = element_rect(fill=NA),
                    panel.grid.major = element_blank(),
                    panel.grid.minor = element_blank(),
                    panel.border = element_rect(fill=NA,color="black", size=.8,
                                                linetype="solid"),
                    axis.title.x=element_text(vjust=-.6, size=18),
                    axis.title.y=element_text(face="bold",vjust=1, size=18),
                    axis.text=element_text(color="black", size=16),
                    legend.title=element_blank(),
                    legend.text=element_text(size=16), 
                    legend.position="bottom",
                    legend.background=element_blank())

pl.pattern <- ggplot( data=df.below.thresh.pattern ) +
  aes( x=as.factor(Channel), y=rms.amp, fill=Pattern) + 
  geom_bar( stat="identity", width=0.8, position=dodge ) +
  geom_errorbar( limits, position=dodge, width=0.15 ) +
  xlab("Channel") +
  ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
  pl.theme.bar
pl.pattern
```

#### `r paste( this.harm, " Channels Below p<.0005 for Speed", sep="")`

```{r figure-1F1-speed-ampl-adult}
df.below.thresh.speed <- df.below.thresh %>%
    group_by( Channel, Speed, iSess ) %>%
    summarise( sr.sub.mean=mean(Sr), 
               si.sub.mean=mean(Si),
               rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
    group_by( Channel, Speed ) %>%
    summarise( sr.group.mean=mean( sr.sub.mean ),
               si.group.mean=mean( si.sub.mean ),
               sr.group.sd=sd(sr.sub.mean),
               si.group.sd=sd(si.sub.mean),
               rms.amp=sqrt( sr.group.mean^2 + si.group.mean^2),
               rms.amp.sem = mean( rms.amp.sub )/sqrt(n()),
               nsubs=n(),
               sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
               si.group.sem=mean(si.sub.mean)/sqrt(n())
    )
  
pl.speed <- ggplot( data=df.below.thresh.speed ) +
  aes( x=as.factor(Channel), y=rms.amp, fill=Speed) + 
  geom_bar( stat="identity", width=0.8, position=dodge ) +
  geom_errorbar( limits, position=dodge, width=0.15 ) +
  xlab("Speed") +
  ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
  pl.theme.bar +
  annotate("rect", xmin = 9.5, xmax = 11.5, ymin = 0, ymax = 1.0, alpha = .1)
  
pl.speed
```

#### Save by-subject data for summaries

Do for Pattern then Speed. Ignore interaction because there are no effects.

##### Pattern
```{r}
# Summarize by subject first
df.below.thresh.pattern.bysub <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == "1F1" ) %>%
  group_by( iSess, Channel, Pattern ) %>%
  summarise( sr.mean.bysub=mean(Sr), 
             si.mean.bysub=mean(Si),
             sr.sem=sd(Sr)/sqrt(n()),
             si.sem=sd(Si)/sqrt(n())           
  ) 

# Then, summarize averages across subjects
df.below.thresh.pattern.summ <- df.below.thresh.pattern.bysub %>%
  group_by( Channel, Pattern ) %>%
  summarise( sr.mean = mean( sr.mean.bysub ),
             si.mean = mean( si.mean.bysub ),
             sr.sem = sd(sr.mean.bysub)/sqrt(n()),
             si.sem = sd(si.mean.bysub)/sqrt(n())
  )
```

##### Speed

```{r}
df.below.thresh.speed.bysub <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm ) %>%
  group_by( iSess, Channel, Speed ) %>%
  summarise( sr.mean.bysub=mean(Sr), 
             si.mean.bysub=mean(Si),
             sr.sem=sd(Sr)/sqrt(n()),
             si.sem=sd(Si)/sqrt(n())           
  ) 

df.below.thresh.speed.summ <- df.below.thresh.speed.bysub %>%
  group_by( Channel, Speed ) %>%
  summarise( sr.mean = mean( sr.mean.bysub ),
             si.mean = mean( si.mean.bysub ),
             sr.sem = sd(sr.mean.bysub)/sqrt(n()),
             si.sem = sd(si.mean.bysub)/sqrt(n())
  )
```

#### Figure -- `r paste(this.harm, " Selected Channels below p<.0005 for Pattern", sep="")`

```{r figure-1F1-pattern-selected}
pl.theme.vect <- theme(plot.title = element_text(lineheight=.8, face ="bold", vjust=2, size = 20),
                  axis.title.x=element_text(vjust=-.6, size=18),
                  axis.title.y=element_text(face="bold",vjust=1, size=18),
                  axis.text=element_text(color="black", size=14),
                  legend.title=element_blank(),
                  legend.text=element_text(size=16),
                  legend.background=element_blank(),
                  legend.position="bottom",
                  legend.title = element_blank(),
                  strip.text = element_text( size = 14 ))

df.below.thresh.pattern.summ %>% filter( Channel %in% c(106,107) ) %>%
  ggplot() +
  aes( x=sr.mean, y=si.mean, color=Pattern ) +
  geom_point() +
  geom_segment(aes( xend=0, yend=0, x=sr.mean, y=si.mean)) +
  geom_pointrange(aes(ymin=si.mean-si.sem, ymax=si.mean+si.sem)) +
  geom_errorbarh( aes(xmin=sr.mean-sr.sem, xmax=sr.mean+sr.sem), height=0) +
  scale_x_continuous(limits = c(-.6,.6)) +
  scale_y_continuous(limits = c(-.6,.6)) +
  coord_fixed( ratio=1 ) +
  xlab( expression(paste("Signal Real (", mu, "V)", sep=""))) +
  ylab( expression(paste("Signal Imaginary (", mu, "V)", sep=""))) +
  facet_grid(. ~ Channel) +
  pl.theme.vect
```

#### `r paste(this.harm, " All Channels Below p<.0005", sep="")

```{r figure-1F1-pattern-vector-adult}
pl.theme.vect.all <- theme(axis.title.x=element_text(vjust=-.6, size=18),
                  axis.title.y=element_text(face="bold",vjust=1, size=18),
                  axis.text=element_text(color="black", size=8),
                  legend.title=element_blank(),
                  legend.text=element_text(size=16),
                  legend.background=element_blank(),
                  legend.position="bottom",
                  legend.title = element_blank(),
                  strip.text = element_text( size = 14 ))

amp.max = 1.4
df.below.thresh.pattern.summ %>% 
  ggplot() +
  aes( x=sr.mean, y=si.mean, color=Pattern ) +
  geom_point() +
  geom_segment(aes( xend=0, yend=0, x=sr.mean, y=si.mean)) +
  geom_pointrange(aes(ymin=si.mean-si.sem, ymax=si.mean+si.sem)) +
  geom_errorbarh( aes(xmin=sr.mean-sr.sem, xmax=sr.mean+sr.sem), height=0) +
  coord_fixed( ratio=1 ) +
  xlab( expression(paste("Signal Real (", mu, "V)", sep=""))) +
  ylab( expression(paste("Signal Imaginary (", mu, "V)", sep=""))) +
  facet_wrap(facets= ~ Channel, scales="free") +
  pl.theme.vect.all
```

```{r}
this.harm="1F2"
```

## Mass univariate MANOVA on `r this.harm`

MANOVA on `r this.harm` with Pattern, Speed, and Pattern x Speed. Analyze and write write output to file.

```{r 1F2-mass-univariate-manova}
this.harm = "1F2"

maov.list = lapply( unique( df.moco$Channel ), mass.univariate, df=df.moco, harm=this.harm, method="manova.ampl.complex")
maov.summ = lapply( maov.list, summary )
```

#### Plot channel-level effects

Extract $p$ values from MANOVA summary. Do not adjust for multiple comparisons here as we use our own multiple comparison adjustment procedure. Create data frame for plotting and tweak factor levels and labels.

```{r}  
# Select pvals from MANOVA data
list.pvals = t( sapply(maov.summ, moco.pvals ) )

# Extract p values and create new data frame for plotting. Use unadjusted p values here, as we use our own
  
adj.Pvals = c(p.adjust(list.pvals[,1], method="none"), p.adjust(list.pvals[,2], method="none"), p.adjust(list.pvals[,3], method="none"))
  
# Create data frame from pvals for plotting
df.pvals = data.frame( Chan = rep( 1:128, 3), 
                       Cond = rep(c("Pattern", "Speed", "Patt*Spd"), c(128, 128, 128)),
                         Pvals = as.numeric( adj.Pvals ),
                         xpos=rep(df.egi$xpos,3), 
                         ypos=rep(df.egi$ypos,3)
                         
  )
  
pvals.cuts = c(-.01,.0001, .0005, .001, .005, .01, .05, 1)
pvals.lbls = c("<.0001","<.0005", "<.001", "<.005", "<.01", "<.05", "ns")

# Create cuts based on p-value levels
Pvals.cuts = cut( df.pvals$Pvals, breaks=pvals.cuts, labels=pvals.lbls)
Pvals.cuts = ordered( Pvals.cuts, levels = rev( pvals.lbls ) )
df.pvals$Pvals.cuts = Pvals.cuts

df.pvals$Cond = ordered( df.pvals$Cond, levels=c("Pattern", "Speed", "Patt*Spd"))
df.pvals$Harm = rep( this.harm, 3*128 )
```

#### Figure -- `r paste(this.harm, " Effects by Channel", sep="")`

```{r figure-1F2-effects-by-channel}
pl.title <- paste(this.harm, "Effects by Channel", sep=" ")

# Plot
pl <- ggplot(data=df.pvals, aes(x=xpos, y=ypos, color=Pvals.cuts, size=Pvals.cuts)) +
  geom_point() +
  facet_grid(facets = . ~ Cond) +
  coord_fixed() + 
  xquiet + 
  yquiet + 
  theme( plot.title = element_text(lineheight=.8, face ="bold", vjust=2, size=16),
         legend.title=element_text(size=0),
         legend.text=element_text(size=12, face="bold"),
         legend.position="bottom") 

# Add ears, nose
pl + annotation_custom(topo_ears_nose, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)
```

Notice midline cluster of channels.

#### Extract data for channels below threshold

```{r}
chans.below.thresh = df.pvals$Chan[df.pvals$Pvals <= p.thresh ]

df.below.thresh <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm )
```

#### Pattern

Calculate mean values for each participant, by channel and pattern for the real (Sr) and imaginary (Si) components at the selected harmonic. Then, calculate the group means for Sr and Si and the group standard deviations and root-mean-square (RMS) amplitudes (and standard errors) from the group mean Sr and Si values.

```{r}
df.below.thresh.pattern <- df.below.thresh %>%
  group_by( Channel, Pattern, iSess ) %>%
  summarise( sr.sub.mean=mean(Sr), 
             si.sub.mean=mean(Si),
             rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
  group_by( Channel, Pattern ) %>%
  summarise( sr.group.mean=mean( sr.sub.mean ),
             si.group.mean=mean( si.sub.mean ),
             sr.group.sd=sd(sr.sub.mean),
             si.group.sd=sd(si.sub.mean),
             rms.amp=sqrt( sr.group.mean^2 + si.group.mean^2),
             rms.amp.sem = mean( rms.amp.sub )/sqrt(n()),
             nsubs=n(),
             sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
             si.group.sem=mean(si.sub.mean)/sqrt(n())
             )
```

#### `r paste( this.harm, " Channels Below p<.0005 for Pattern", sep="")`

```{r figure-1F2-pattern-ampl}
limits <- aes( ymax = rms.amp + rms.amp.sem, ymin = rms.amp )
dodge <- position_dodge( width=0.8 )

pl.pattern <- ggplot( data=df.below.thresh.pattern ) +
  aes( x=as.factor(Channel), y=rms.amp, fill=Pattern) + 
  geom_bar( stat="identity", width=0.8, position=dodge ) +
  geom_errorbar( limits, position=dodge, width=0.15 ) +
  xlab("Channel") +
  ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
  pl.theme.bar +
  annotate("rect", xmin = 5.5, xmax = 6.5, ymin = 0, ymax = 0.5, alpha = .1) + # Chan 62
  annotate("rect", xmin = 15-.5, xmax = 15+.5, ymin = 0, ymax = 0.5, alpha = .1) # Chan 75

pl.pattern
```

#### Figure -- `r paste( this.harm, " Channels Below p<.0005 for Speed", sep="")`

```{r figure-1F2-speed-ampl}
df.below.thresh.speed <- df.below.thresh %>%
    group_by( Channel, Speed, iSess ) %>%
    summarise( sr.sub.mean=mean(Sr), 
               si.sub.mean=mean(Si),
               rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
    group_by( Channel, Speed ) %>%
    summarise( sr.group.mean=mean( sr.sub.mean ),
               si.group.mean=mean( si.sub.mean ),
               sr.group.sd=sd(sr.sub.mean),
               si.group.sd=sd(si.sub.mean),
               rms.amp=sqrt( sr.group.mean^2 + si.group.mean^2),
               rms.amp.sem = mean( rms.amp.sub )/sqrt(n()),
               nsubs=n(),
               sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
               si.group.sem=mean(si.sub.mean)/sqrt(n())
    )
  
pl.speed <- ggplot( data=df.below.thresh.speed ) +
  aes( x=as.factor(Channel), y=rms.amp, fill=Speed) + 
  geom_bar( stat="identity", width=0.8, position=dodge ) +
  geom_errorbar( limits, position=dodge, width=0.15 ) +
  xlab("Speed") +
  ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
  pl.theme.bar +
  annotate("rect", xmin = 5.5, xmax = 6.5, ymin = 0, ymax = 0.5, alpha = .1) + # Chan 62
  annotate("rect", xmin = 15-.5, xmax = 15+.5, ymin = 0, ymax = 0.5, alpha = .1) # Chan 75
  
pl.speed
```

#### Save by-subject data for summaries

Do for Pattern then Speed. Ignore interaction because there are no effects.

##### Pattern
```{r figure-1F2-select-chans-below-pattern}
# Summarize by subject first
df.below.thresh.pattern.bysub <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == "1F1" ) %>%
  group_by( iSess, Channel, Pattern ) %>%
  summarise( sr.mean.bysub=mean(Sr), 
             si.mean.bysub=mean(Si),
             sr.sem=sd(Sr)/sqrt(n()),
             si.sem=sd(Si)/sqrt(n())           
  ) 

# Then, summarize averages across subjects
df.below.thresh.pattern.summ <- df.below.thresh.pattern.bysub %>%
  group_by( Channel, Pattern ) %>%
  summarise( sr.mean = mean( sr.mean.bysub ),
             si.mean = mean( si.mean.bysub ),
             sr.sem = sd(sr.mean.bysub)/sqrt(n()),
             si.sem = sd(si.mean.bysub)/sqrt(n())
  )
```

##### Speed

```{r 1F2-select-chans-below-spd}
df.below.thresh.speed.bysub <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm ) %>%
  group_by( iSess, Channel, Speed ) %>%
  summarise( sr.mean.bysub=mean(Sr), 
             si.mean.bysub=mean(Si),
             sr.sem=sd(Sr)/sqrt(n()),
             si.sem=sd(Si)/sqrt(n())           
  ) 

df.below.thresh.speed.summ <- df.below.thresh.speed.bysub %>%
  group_by( Channel, Speed ) %>%
  summarise( sr.mean = mean( sr.mean.bysub ),
             si.mean = mean( si.mean.bysub ),
             sr.sem = sd(sr.mean.bysub)/sqrt(n()),
             si.sem = sd(si.mean.bysub)/sqrt(n())
  )
```

#### Figure -- `r paste(this.harm, " Selected Channels below p<.0005 for Speed", sep="")`

Select channels 62 and 75.

```{r figure-1F2-speed-selected}
df.below.thresh.speed.summ %>% filter( Channel %in% c(62,75) ) %>%
  ggplot() +
  aes( x=sr.mean, y=si.mean, color=Speed ) +
  geom_point() +
  geom_segment(aes( xend=0, yend=0, x=sr.mean, y=si.mean)) +
  geom_pointrange(aes(ymin=si.mean-si.sem, ymax=si.mean+si.sem)) +
  geom_errorbarh( aes(xmin=sr.mean-sr.sem, xmax=sr.mean+sr.sem), height=0) +
  scale_x_continuous(limits = c(-.6,.6)) +
  scale_y_continuous(limits = c(-.6,.6)) +
  coord_fixed( ratio=1 ) +
  xlab( expression(paste("Signal Real (", mu, "V)", sep=""))) +
  ylab( expression(paste("Signal Imaginary (", mu, "V)", sep=""))) +
  facet_grid(. ~ Channel) +
  pl.theme.vect
```

#### `r paste(this.harm, " All Channels Below p<.0005 for Speed", sep="")`

```{r figure-1F2-speed-vector}
amp.max = 0.5
df.below.thresh.speed.summ %>% 
  ggplot() +
  aes( x=sr.mean, y=si.mean, color=Speed ) +
  geom_point() +
  geom_segment(aes( xend=0, yend=0, x=sr.mean, y=si.mean)) +
  geom_pointrange(aes(ymin=si.mean-si.sem, ymax=si.mean+si.sem)) +
  geom_errorbarh(aes(xmin=sr.mean-sr.sem, xmax=sr.mean+sr.sem), height=0) +
  coord_fixed( ratio=1 ) +
  xlab( expression(paste("Signal Real (", mu, "V)", sep=""))) +
  ylab( expression(paste("Signal Imaginary (", mu, "V)", sep=""))) +
  facet_wrap(facets= ~ Channel, scales = "free") +
  pl.theme.vect.all
```

```{r}
this.harm="2F1"
```

## Mass univariate MANOVA on `r this.harm`

MANOVA on `r this.harm` with Pattern, Speed, and Pattern x Speed. Analyze and write write output to file.

```{r 2F1-mass-univariate-manova}
maov.list = lapply( unique( df.moco$Channel ), mass.univariate, df=df.moco, harm=this.harm, method="manova.ampl.complex")
maov.summ = lapply( maov.list, summary )
```

#### Plot channel-level effects

Extract $p$ values from MANOVA summary. Do not adjust for multiple comparisons here as we use our own multiple comparison adjustment procedure. Create data frame for plotting and tweak factor levels and labels.

```{r}  
# Select pvals from MANOVA data
list.pvals = t( sapply(maov.summ, moco.pvals ) )

# Extract p values and create new data frame for plotting. Use unadjusted p values here, as we use our own
  
adj.Pvals = c(p.adjust(list.pvals[,1], method="none"), p.adjust(list.pvals[,2], method="none"), p.adjust(list.pvals[,3], method="none"))
  
# Create data frame from pvals for plotting
df.pvals = data.frame( Chan = rep( 1:128, 3), 
                       Cond = rep(c("Pattern", "Speed", "Patt*Spd"), c(128, 128, 128)),
                         Pvals = as.numeric( adj.Pvals ),
                         xpos=rep(df.egi$xpos,3), 
                         ypos=rep(df.egi$ypos,3)
                         
  )
  
pvals.cuts = c(-.01,.0001, .0005, .001, .005, .01, .05, 1)
pvals.lbls = c("<.0001","<.0005", "<.001", "<.005", "<.01", "<.05", "ns")

# Create cuts based on p-value levels
Pvals.cuts = cut( df.pvals$Pvals, breaks=pvals.cuts, labels=pvals.lbls)
Pvals.cuts = ordered( Pvals.cuts, levels = rev( pvals.lbls ) )
df.pvals$Pvals.cuts = Pvals.cuts

df.pvals$Cond = ordered( df.pvals$Cond, levels=c("Pattern", "Speed", "Patt*Spd"))
df.pvals$Harm = rep( this.harm, 3*128 )
```

#### `r paste(this.harm, " Effects By Channel", sep="")`

```{r figure-2F1-effects-by-channel-adult}
pl.title <- paste(this.harm, "Effects by Channel", sep=" ")

# Plot
pl <- ggplot(data=df.pvals, aes(x=xpos, y=ypos, color=Pvals.cuts, size=Pvals.cuts)) +
  geom_point() +
  facet_grid(facets = . ~ Cond) +
  coord_fixed() + 
  xquiet + 
  yquiet + 
  pl.theme.topo

# Add ears, nose
pl + annotation_custom(topo_ears_nose, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)
```

No channels meet $p$<.0005 so we ignore.

```{r}
this.harm = "3F1"
```

## Mass univariate MANOVA on `r this.harm`

MANOVA on `r this.harm` with Pattern, Speed, and Pattern x Speed. 

```{r 3F1-mass-univariate-manova}
maov.list = lapply( unique( df.moco$Channel ), mass.univariate, df=df.moco, harm=this.harm, method="manova.ampl.complex")
maov.summ = lapply( maov.list, summary )
```

#### Plot channel-level effects

Extract $p$ values from MANOVA summary. Do not adjust for multiple comparisons here as we use our own multiple comparison adjustment procedure. Create data frame for plotting and tweak factor levels and labels.

```{r}  
# Select pvals from MANOVA data
list.pvals = t( sapply(maov.summ, moco.pvals ) )

# Extract p values and create new data frame for plotting. Use unadjusted p values here, as we use our own
  
adj.Pvals = c(p.adjust(list.pvals[,1], method="none"), p.adjust(list.pvals[,2], method="none"), p.adjust(list.pvals[,3], method="none"))
  
# Create data frame from pvals for plotting
df.pvals = data.frame( Chan = rep( 1:128, 3), 
                       Cond = rep(c("Pattern", "Speed", "Patt*Spd"), c(128, 128, 128)),
                         Pvals = as.numeric( adj.Pvals ),
                         xpos=rep(df.egi$xpos,3), 
                         ypos=rep(df.egi$ypos,3)
                         
  )
  
pvals.cuts = c(-.01,.0001, .0005, .001, .005, .01, .05, 1)
pvals.lbls = c("<.0001","<.0005", "<.001", "<.005", "<.01", "<.05", "ns")

# Create cuts based on p-value levels
Pvals.cuts = cut( df.pvals$Pvals, breaks=pvals.cuts, labels=pvals.lbls)
Pvals.cuts = ordered( Pvals.cuts, levels = rev( pvals.lbls ) )
df.pvals$Pvals.cuts = Pvals.cuts

df.pvals$Cond = ordered( df.pvals$Cond, levels=c("Pattern", "Speed", "Patt*Spd"))
df.pvals$Harm = rep( this.harm, 3*128 )
```

#### `r paste(this.harm, " Effects By Channel", sep="")`

```{r figure-3F1-effects-by-channel-adult}
pl.title <- paste(this.harm, "Effects by Channel", sep=" ")

# Plot
pl <- ggplot(data=df.pvals, aes(x=xpos, y=ypos, color=Pvals.cuts, size=Pvals.cuts)) +
  geom_point() +
  facet_grid(facets = . ~ Cond) +
  coord_fixed() + 
  xquiet + 
  yquiet + 
  pl.theme.topo

# Add ears, nose
pl + annotation_custom(topo_ears_nose, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf)
```

Interesting midline cluster at p<.0001 and p<.0005 for speed.

#### Extract data for channels below threshold

```{r}
chans.below.thresh = df.pvals$Chan[df.pvals$Pvals <= p.thresh ]

df.below.thresh <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm )
```

#### `r paste( this.harm, " Channels Below p<.0005 for Speed", sep="")`

```{r figure-3F1-speed-ampl-adult}
df.below.thresh.speed <- df.below.thresh %>%
    group_by( Channel, Speed, iSess ) %>%
    summarise( sr.sub.mean=mean(Sr), 
               si.sub.mean=mean(Si),
               rms.amp.sub=sqrt(sr.sub.mean^2+si.sub.mean^2)) %>%
    group_by( Channel, Speed ) %>%
    summarise( sr.group.mean=mean( sr.sub.mean ),
               si.group.mean=mean( si.sub.mean ),
               sr.group.sd=sd(sr.sub.mean),
               si.group.sd=sd(si.sub.mean),
               rms.amp=sqrt( sr.group.mean^2 + si.group.mean^2),
               rms.amp.sem = mean( rms.amp.sub )/sqrt(n()),
               nsubs=n(),
               sr.group.sem=mean(sr.sub.mean)/sqrt(n()),
               si.group.sem=mean(si.sub.mean)/sqrt(n())
    )
  
pl.speed <- ggplot( data=df.below.thresh.speed ) +
  aes( x=as.factor(Channel), y=rms.amp, fill=Speed) + 
  geom_bar( stat="identity", width=0.8, position=dodge ) +
  geom_errorbar( limits, position=dodge, width=0.15 ) +
  xlab("Speed") +
  ylab(expression(paste("RMS amplitude (", mu, "V)", sep=""))) +
  pl.theme.bar
  
pl.speed
```

#### Save by-subject data for summaries

Do for Pattern then Speed. Ignore interaction because there are no effects.

##### Pattern

```{r 3F1-select-chans-below-patt}
# Summarize by subject first
df.below.thresh.pattern.bysub <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == "1F1" ) %>%
  group_by( iSess, Channel, Pattern ) %>%
  summarise( sr.mean.bysub=mean(Sr), 
             si.mean.bysub=mean(Si),
             sr.sem=sd(Sr)/sqrt(n()),
             si.sem=sd(Si)/sqrt(n())           
  ) 

df.below.thresh.pattern.summ <- df.below.thresh.pattern.bysub %>%
  group_by( Channel, Pattern ) %>%
  summarise( sr.mean = mean( sr.mean.bysub ),
             si.mean = mean( si.mean.bysub ),
             sr.sem = sd(sr.mean.bysub)/sqrt(n()),
             si.sem = sd(si.mean.bysub)/sqrt(n())
  )
```

##### Speed

```{r 3F1-select-chans-below-spd}
df.below.thresh.speed.bysub <- df.moco %>%
  filter( Channel %in% chans.below.thresh, Harm == this.harm ) %>%
  group_by( iSess, Channel, Speed ) %>%
  summarise( sr.mean.bysub=mean(Sr), 
             si.mean.bysub=mean(Si),
             sr.sem=sd(Sr)/sqrt(n()),
             si.sem=sd(Si)/sqrt(n())           
  ) 

df.below.thresh.speed.summ <- df.below.thresh.speed.bysub %>%
  group_by( Channel, Speed ) %>%
  summarise( sr.mean = mean( sr.mean.bysub ),
             si.mean = mean( si.mean.bysub ),
             sr.sem = sd(sr.mean.bysub)/sqrt(n()),
             si.sem = sd(si.mean.bysub)/sqrt(n())
  )
```

#### `r paste(this.harm, " All Channels Below p<.0005 for Speed", sep="")`

```{r figure-3F1-speed-vector-adult}
amp.max = 0.5
df.below.thresh.speed.summ %>% 
  ggplot() +
  aes( x=sr.mean, y=si.mean, color=Speed ) +
  geom_point() +
  geom_segment(aes( xend=0, yend=0, x=sr.mean, y=si.mean)) +
  geom_pointrange(aes(ymin=si.mean-si.sem, ymax=si.mean+si.sem)) +
  geom_errorbarh( aes(xmin=sr.mean-sr.sem, xmax=sr.mean+sr.sem), height=0) +
  coord_fixed( ratio=1 ) +
  xlab( expression(paste("Signal Real (", mu, "V)", sep=""))) +
  ylab( expression(paste("Signal Imaginary (", mu, "V)", sep=""))) +
  facet_wrap(facets= ~ Channel, scales = "free") +
  pl.theme.vect.all

```

Responses to 4 and 8 deg/s have more similar phase tuning profiles. In fact, phase matters a great deal for channels 84 and 89 where amplitudes look similar.

